---
title: "20220707_fpl"
author: "Arif P. Sulistiono / @arifpras"
format: html
editor: visual
---

# Prerequisites

This initial step covers clearing the environment, installing/loading related libraries that will be used through the analysis, and setting the working directory.

```{r pre-requisites}
# Remove all items in Environment!
rm(list = ls())
ls()

# Load packages
# pacman::p_load(tidyverse, data.table, readxl, writexl, glue, wesanderson,
#                RColorBrewer, viridis, ggrepel, ggthemes, ggsci, mFilter,
#                quantmod, corrplot, cowplot, readstata13, readr, knitr,
#                kableExtra, stargazer, parameters, colorspace, dendextend,
#                ggdendro, purrr, cluster, plotly, DescTools, factoextra,
#                grid, gridExtra, ggsoccer, RCurl, tidytext, rvest)

pacman::p_load(tidyverse,
               stats,
               ggsoccer,
               utils,
               tidytext,
               rvest,
               corrr,
               jsonlite,
               fuzzyjoin,
               stringr,
               viridis)

options(tz = "Europe/London")

# Working directory
setwd("~/OneDrive - The University of Nottingham/BB_SideProject/BelutListrik")
```

# Building database

## FPL: 2020/2021 and 2021/2022

From the FPL's database, we want to get information on players' performances in each game week in two seasons. To obtain the information, we will perform 'web-scrapping' on following repositories: "<https://github.com/vaastav/Fantasy-Premier-League/tree/master/data/2020-21/gws>" and "<https://github.com/vaastav/Fantasy-Premier-League/tree/master/data/2021-22/gws>". *Note: download the RAW version.*

```{r db_fpl}
# Season 2020/2021

fpl20 <-
  "https://raw.githubusercontent.com/vaastav/Fantasy-Premier-League/master/data/2020-21/gws/merged_gw.csv"

(fpl01 <- read_csv(fpl20, show_col_types = FALSE) %>%
    mutate(season = "20/21"))

# Season 2021/2022

fpl21 <-
  "https://raw.githubusercontent.com/vaastav/Fantasy-Premier-League/master/data/2021-22/gws/merged_gw.csv"

(fpl02 <- read_csv(fpl21, show_col_types = FALSE) %>%
    mutate(season = "21/22"))

# Combine those seasons

(fpl03 <- fpl01 %>%
    bind_rows(fpl02))
```

## Aggregating FPL's statistics

The objective is to understand the average points, minutes, ICT index, threats, creativity, and influence in the last two seasons.

```{r fpl_agg}
(
  fpl04 <- fpl03 %>%
    group_by(name) %>%
    summarise(
      avg_points = mean(total_points),
      avg_minutes = mean(minutes),
      avg_ict = mean(ict_index),
      avg_threat = mean(threat),
      avg_creativity = mean(creativity),
      avg_influence = mean(influence),
      .groups = "keep"
    ) %>%
    ungroup() %>%
    arrange(desc(avg_points))
)

fpl04 %>% select(-name) %>%
  tibble() %>%
  correlate() %>%
  network_plot(min_cor = 0.9)

# Setting more restriction:

# (fpl04 %>%  filter(avg_points > 2 & avg_minutes > 40))
```

## Understat: 2020/2021 and 2021/2022

From the Understat's database, we want to get other information (e.g., xG, xA, shots, key passes, npg, npxG, xGChain, xGBuildUp). The 'web-scrapping' will be performed on following repositories: "<https://github.com/vaastav/Fantasy-Premier-League/tree/master/data/2020-21>" and "<https://github.com/vaastav/Fantasy-Premier-League/tree/master/data/2021-22/understat>". *Note: download the RAW version.*

```{r db_understat}
# Season 2020/2021

uds20 <-
  "https://raw.githubusercontent.com/arifpras/Fantasy-Premier-League/master/data/2020-21/understat/understat_player.csv"

(
  uds01 <- read_csv(uds20, show_col_types = FALSE) %>%
    mutate(
      season = "20/21",
      xG90 = xG / time * 90,
      xA90 = xA / time * 90,
      npxG90 = npxG / time * 90,
      xGChain90 = xGChain / time * 90,
      xGBuildup90 = xGBuildup / time * 90
    )
)

# Season 2021/2022

uds21 <-
  "https://raw.githubusercontent.com/arifpras/Fantasy-Premier-League/master/data/2021-22/understat/understat_player.csv"


(
  uds02 <- read_csv(uds21, show_col_types = FALSE) %>%
    mutate(
      season = "21/22",
      xG90 = xG / time * 90,
      xA90 = xA / time * 90,
      npxG90 = npxG / time * 90,
      xGChain90 = xGChain / time * 90,
      xGBuildup90 = xGBuildup / time * 90
    )
)

# Combine those seasons

(uds03 <- uds01 %>%
    bind_rows(uds02))
```

## Aggregating Understat's statistics

The objective is to understand the average xG (expected goals), xA (expected assist), npxG (non-penalty expected goals), xGChain (number of Expected Goals in which a player has participated), and xGBuildup (which players are part of passing chains that end in a shot but without taking into account the last two links -- shot and last pass -, thus measuring midfielders, centre-backs or full-backs who have an important influence on their team's possessions) in the last two seasons.

```{r understat_agg}
(
  uds04 <- uds03 %>%
    group_by(player_name, id) %>%
    summarise(
      avg_xG = mean(xG),
      #expected goals
      avg_xA = mean(xA),
      #expected assist
      avg_npxG = mean(npxG),
      #non-penalty expected goals
      avg_xGChain = mean(xGChain),
      #number of Expected Goals in which a player has participated
      avg_xGBuildup = mean(xGBuildup),
      #which players are part of passing chains that end in a shot but without taking into account the last two links – shot and last pass -, thus measuring midfielders, centre-backs or full-backs who have an important influence on their team's possessions,
      avg_xG90 = mean(xG90),
      #every 90 minutes
      avg_xA90 = mean(xA90),
      #every 90 minutes
      avg_npxG90 = mean(npxG90),
      #every 90 minutes
      avg_xGChain90 = mean(xGChain90),
      #every 90 minutes
      avg_xGBuildup90 = mean(xGBuildup90),
      #every 90 minutes
      .groups = "keep"
    ) %>%
    ungroup() %>%
    arrange(desc(avg_xGChain))
)

uds04 %>%
  select(avg_xG, avg_xA, avg_npxG, avg_xGChain, avg_xGBuildup) %>%
  tibble() %>%
  correlate() %>%
  network_plot(min_cor = 0.7)
```

## Merging the databases

To do so, we need to have another file. Another web-scrapping needs to be done.

```{r dictionary}
dict01 <-
  "https://raw.githubusercontent.com/arifpras/Fantasy-Premier-League/master/data/2021-22/id_dict.csv"

(dict02 <- readr::read_csv(dict01, show_col_types = FALSE))

(dict03 <- fpl01 %>% select(name, position) %>% distinct(name, position))
(dict04 <- fpl02 %>% select(name, position) %>% distinct(name, position))

(dict05 <- dict03 %>%
    bind_rows(dict04) %>%
    arrange(name) %>%
    distinct(name, position) %>%
    right_join(
      dict02, by = c("name" = "FPL_Name")
    ))

# Check duplicated players
dict05 %>% group_by(name) %>% filter(n()>1)

# Stuart Dallas: DEF
# Matt Ritchie: DEF
# Allan Saint-Maximin: MID
# Ainsley Maitland-Niles: MID

(dict06 <- dict05 %>%
    filter(!((name == "Stuart Dallas" &
                position != "MID") |
               (name == "Matt Ritchie" &
                  position != "MID") |
               (name == "Allan Saint-Maximin" & position != "FWD") |
               (name == "Ainsley Maitland-Niles" &
                  position != "DEF") |
               (
                 name %in% c(
                   "Álvaro Fernández",
                   "Cheikhou Kouyaté",
                   "Fernando Luiz Rosa",
                   "Nathan Broadhead",
                   "Pierre-Emerick Aubameyang"
                 )
               )
    )))

# Check duplicated players
dict06 %>% group_by(name) %>% filter(n()>1)

# Not in the EPL anymore
dict02 %>% anti_join(dict06, by = c("Understat_Name"))
```

Now, join all the databases.

```{r db01}
# FPL data and the dictionary

(
  fpl05 <- fpl04 %>%
    select(name, avg_points, avg_minutes, avg_ict, avg_influence) %>%
    right_join(dict02, by = c("name" = "FPL_Name")) %>%
    select(-Understat_ID, -name, -Understat_Name)
)

# Understat data and the dictionary

(
  uds05 <- uds04 %>%
    select(player_name, id, avg_xG, avg_xA, avg_xGChain) %>%
    right_join(dict02, by = c("id" = "Understat_ID")) %>%
    select(-player_name) %>%
    rename(
      Understat_ID = id
    ) %>%
    relocate(
      Understat_Name, Understat_ID, FPL_Name, FPL_ID
    )
)

# Joining fpl05 and uds05

(
  db01 <- fpl05 %>%
    left_join(uds05, by = c("FPL_ID")) %>%
    relocate(
      Understat_Name, Understat_ID, FPL_Name, FPL_ID
    )
)
```

## 2022/2023's prices

The objective is to find prices for all players in the upcoming season. We will do another web-scrapping on the FPL official database.

```{r db_official}
(off00 <-
   fromJSON("https://fantasy.premierleague.com/api/bootstrap-static/"))
```

By exploring the JSON data set, we can obtain the prices.

```{r db_prices}
(
  pr01 <- tibble(off00$elements) %>%
    select(first_name,
           second_name,
           web_name,
           now_cost,
           team,
           team_code,
           #id,
           element_type,
           # points_per_game,
           # selected_by_percent,total_points,
           # value_form,
           # value_season,
           # minutes,
           # goals_scored,
           # assists,
           # clean_sheets,
           # goals_conceded,
           # saves,
           # bps,
           # influence,
           # creativity,
           # threat,
           # ict_index,
           # influence_rank,
           # influence_rank_type,
           # creativity_rank,
           # creativity_rank_type,
           # threat_rank,
           # threat_rank_type,
           # ict_index_rank,
           # ict_index_rank_type,

           ) %>%
    mutate(now_cost = now_cost / 10)
    )
```

...we also can get information about the strength of each team (difficulty rating).

```{r team}
#team
(tm01 <- tibble(off00$teams) %>%
  select(id, name, short_name, strength))
```

...and the position for each player.

```{r position}
(
  pos01 <- tibble(off00$element_types) %>%
    select(id, singular_name, singular_name_short)
)
```

Now, we can merge all information from the official website.

```{r db02}
(
  db02 <- pr01 %>%
    left_join(tm01, by = c("team" = "id")) %>%
    left_join(pos01, by = c("element_type" = "id")) %>%
    rename(
      "team_name" = name,
      "position_name" = singular_name,
      "team_strength" = strength
    )  %>%
    #str()
    mutate_if(is.integer, as.numeric) %>%
    #str()
    mutate(full_name = paste(first_name, second_name, sep = " ")) %>%
    select(
      -team,
      -team_code,
      -element_type,
      -first_name,
      -second_name
    ) %>%
    relocate(full_name, team_name, team_strength, position_name, now_cost)
)
```

```{r partial_match}
# (
#   db03 <- db02 %>%
#     fuzzy_full_join(
#       dict06,
#       by = c("full_name" = "Understat_Name", "singular_name_short" = "position"),
#       match_fun = stringi::stri_detect_fixed
#     )
# )


# xlsx::write.xlsx(db02, file = "20220708_matching.xlsx", sheetName = "new",
#   col.names = TRUE, row.names = TRUE, append = FALSE)
# xlsx::write.xlsx(dict06, file = "20220708_matching.xlsx", sheetName = "old",
#   col.names = TRUE, row.names = TRUE, append = TRUE)

# Check duplicated players
# (db04 <- db03 %>% group_by(full_name) %>% filter(n()>1))
```

```{r db03}
db03 <-
  "https://raw.githubusercontent.com/arifpras/BelutListrik/main/idlist2223.csv?token=GHSAT0AAAAAABVVEBDOJEQYVTHYHQYW4X6IYWJZD6A"

(db04 <- read_csv(db03, show_col_types = FALSE))
```

Combining all players' statistics.

```{r db05}
(
  db05 <- db04 %>% select(-Understat_Name) %>%
    left_join(db01, by = c("FPL_ID"))
)

str(db05)

# (db06 <- extract(db05, full_name, c("first_name", "last_name"), "([^ ]+) (.*)"))

(
  db06 <- db05 %>%
    left_join(db02 %>% select(full_name, web_name, team_name), by = c("full_name", "team_name")) %>%
    select(-short_name,-name,-FPL_ID,-FPL_Name, -Understat_Name,-Understat_ID) %>%
    relocate(full_name, web_name)
)
```

## Fixture difficulty

First, we need to obtain the upcoming fixtures. Another web-screapping will be done on the following website <https://fixturedownload.com/results/epl-2022>.

```{r fixture}
fixt00 <- "https://fixturedownload.com/results/epl-2022"
fixt01 <- read_html(fixt00)

fixt02 <- fixt01 %>%
  html_node("body") %>%
  html_children()

fixt03 <- html_nodes(fixt02, 'td')

fixt04 <- html_text(fixt03)
fixt05 <- data.frame(fixt04)

(fixt06 <-
    as.data.frame(matrix(
      fixt05$fixt04, ncol = 6, byrow = TRUE
    )))
```

Then, we will combine the fixtures with each team's strength.

```{r strength}
(
  st01 <- fixt06 %>%
    select(V1, V4, V5) %>%
    rename(
      gameweek = V1,
      home = V4,
      away = V5
    ) %>%
    mutate(
      home = recode(home, "Nottingham Forest" = "Nott'm Forest"),
      away = recode(away, "Nottingham Forest" = "Nott'm Forest")
    ) %>%
    left_join(tm01, by = c("away" = "name")) %>%
    select(-id,-short_name) %>%
    rename(diff_away = strength)  %>%
    left_join(tm01, by = c("home" = "name")) %>%
    select(-id,-short_name) %>%
    rename(diff_home = strength)
)

(
  st02 <- st01 %>%
    select(gameweek, home, diff_away) %>%
    rename(team = home,
           opponent_diff = diff_away)
)

(
  st03 <- st01 %>%
    select(gameweek, away, diff_home) %>%
    rename(team = away,
           opponent_diff = diff_home)
)

(st04 <- st02 %>%
    bind_rows(st03) %>%
    arrange(gameweek) %>%
    mutate(
      gameweek = as.numeric(gameweek),
      opponent_diff = as.numeric(opponent_diff)
    )
)

```

Now, we want to aggregate level of difficulty in the next 8 matches (before the World Cup).

```{r diff_9matchs}
(diff01 <- st04 %>%
   arrange(team, gameweek) %>%
   filter(gameweek < 9) %>%
   group_by(team) %>%
   summarise(diff_opponent = mean(opponent_diff)) %>%
   ungroup() %>%
   arrange(diff_opponent))

# st04 %>% filter(gameweek < 10) %>% arrange(team, gameweek)
```

Combining the next 8 matches difficulty to players' statistics.

```{r db07_final}
(
  db07 <- db06 %>%
    left_join(diff01, by = c ("team_name" = "team")) %>%
    mutate(player_id = row_number()) %>%
    relocate(player_id, full_name, web_name) %>%
    left_join(tm01 %>% select(name, short_name), by = c("team_name" = "name")) %>%
    mutate(player_team = paste(web_name, short_name, sep = " | ")) %>%
    relocate(
      player_id,
      player_team,
      position_name,
      now_cost,
      avg_points,
      avg_minutes,
      avg_ict,
      avg_influence,
      avg_xG,
      avg_xA,
      avg_xGChain,
      diff_opponent
    )
)

saveRDS(db07, file = "db07.rds")

```

# Building team

## Grouping by positions

```{r pos01}
(
  pos01 <- db07 %>%
    select(player_team, position_name) %>%
    as_tibble() %>%
    arrange(position_name) %>%
    group_by(position_name) %>%
    mutate(
      pos_id = row_number()
    ) %>%
    ungroup() %>%
    pivot_wider(
      names_from = position_name,
      values_from = player_team
    ) %>%
    select(-pos_id) %>%
    relocate(
      Goalkeeper, Defender, Midfielder, Forward
    )
)
```

## GK: A group with 2 players

First we want to make a group consists of 2 players.

```{r gk02}
gk01 <- db07 %>% filter(position_name %in% "Goalkeeper")
gk02 <-  t(combn(gk01$player_team, 2)) #max: 2 players

saveRDS(gk02, file = "gk02.rds")
```

...then combine with their statistics.

```{r gk03}
(
  gk03 <- gk02 %>%
    # na.omit() %>%
    as_tibble() %>%
    mutate("id_gk" = row_number()) %>%
    relocate(id_gk) %>%
    pivot_longer(
      cols = 2:3,
      names_to = "v",
      values_to = "player_team"
    ) %>%
    select(-v) %>%
    left_join(db07, by = "player_team")
)
```

...aggregate statistics for each group.

```{r gk04}
(
  gk04 <- gk03 %>%
    group_by(id_gk) %>%
    summarise(
      sum_price = sum(now_cost),
      #sum, because we have maximal budget
      avg_avgpoints = mean(avg_points),
      #average, because we have to pick 1 players
      avg_avgminutes = mean(avg_minutes),
      avg_avgict = mean(avg_ict),
      avg_avginfluence = mean(avg_influence),
      avg_avgXG = mean(avg_xG),
      avg_avgxA = mean(avg_xA),
      avg_avgxGChain = mean(avg_xGChain),
      avg_diffopp = mean(diff_opponent),
    ) %>%
    arrange(desc(avg_avgpoints),
            sum_price)
)
```

Let's say the budget for the goalkeeper position is 10.

```{r gk05}
(
  gk05 <- gk04 %>%
    arrange(desc(avg_avgpoints)) %>%
    filter(sum_price < 10 & sum_price > 8) %>%
    slice_head(n = 5) %>%
    left_join(gk03, by = "id_gk") %>%
    relocate(id_gk, player_team)
)
```

...it's going to be easier to plot the data.

```{r gk_plot}
  (
    gk_plot01 <- gk05 %>%
      select(
        id_gk,
        sum_price,
        avg_avgpoints,
        avg_avgminutes,
        avg_avgict,
        avg_avginfluence,
        avg_avgXG,
        avg_avgxA,
        avg_avgxGChain,
        avg_diffopp
      ) %>%
      group_by(id_gk) %>%
      slice_head(n = 1) %>%
      ungroup() %>%
      pivot_longer(
        cols = 2:10,
        names_to = "indicator",
        values_to = "value"
      ) %>%
      ggplot(aes(
        x = reorder_within(id_gk, value, indicator),
        y = value,
        fill = factor(id_gk)
      )) +
      geom_col() +
      #geom_bar(stat = "identity", aes(fill = factor(id_gk))) +
      coord_flip() +
      scale_x_reordered() +
      #scale_fill_viridis(discrete = TRUE) +
      #scale_fill_manual(values = wes_palette("Darjeeling1")) +
      scale_fill_manual(values = c(
        "#ef476f", "#ffd166", "#06d6a0", "#118ab2", "#073b4c"
      )) +
      egg::theme_article() +
      facet_wrap(~ indicator, ncol = 3, scales = "free") +
      theme(legend.position = "none") +
      labs(x = "",
           y = "")
)
```

Which GK team?

```{r gk_choice}
gk_team <- "823"

(gk_choice <-
    gk05 %>% filter(id_gk == gk_team) %>% mutate("position" = "GK"))
```

Take a loot at the player's statistics individually.

```{r gk_indv_stat}
(
  gk_plot02 <- gk_choice %>%
    select(
      player_team,
      now_cost,
      avg_points,
      avg_minutes,
      avg_ict,
      avg_influence,
      avg_xG,
      avg_xA,
      avg_xGChain,
      diff_opponent
    ) %>%
    pivot_longer(
      cols = 2:10,
      names_to = "indicator",
      values_to = "value"
    ) %>%
    ggplot(aes(
      x = reorder_within(player_team, value, indicator),
      y = value,
      fill = factor(player_team)
    )) +
    geom_col() +
    #geom_bar(stat = "identity", aes(fill = factor(id_gk))) +
    coord_flip() +
    scale_x_reordered() +
    scale_fill_viridis(discrete = TRUE) +
    #scale_fill_manual(values = wes_palette("Darjeeling1")) +
    #scale_fill_manual(values = c("#ef476f", "#ffd166")) +
    egg::theme_article() +
    facet_wrap( ~ indicator, ncol = 3, scales = "free") +
    theme(legend.position = "none") +
    labs(x = "",
         y = "")
)
```

## DEF: A group with 5 players

A group consists of 5 players.

```{r def02}
def01 <- db07 %>% filter(position_name %in% "Defender")

#head(def01)
def02 <-  t(combn(def01$player_team, 5)) #max: 5 players
```

...then combine with their statistics.

```{r def03}
(
  gk03 <- gk02 %>%
    # na.omit() %>%
    as_tibble() %>%
    mutate("id_gk" = row_number()) %>%
    relocate(id_gk) %>%
    pivot_longer(
      cols = 2:3,
      names_to = "v",
      values_to = "player_team"
    ) %>%
    select(-v) %>%
    left_join(db07, by = "player_team")
)
```

...aggregate statistics for each group.

```{r def04}
(
  gk04 <- gk03 %>%
    group_by(id_gk) %>%
    summarise(
      sum_price = sum(now_cost),
      #sum, because we have maximal budget
      avg_avgpoints = mean(avg_points),
      #average, because we have to pick 1 players
      avg_avgminutes = mean(avg_minutes),
      avg_avgict = mean(avg_ict),
      avg_avginfluence = mean(avg_influence),
      avg_avgXG = mean(avg_xG),
      avg_avgxA = mean(avg_xA),
      avg_avgxGChain = mean(avg_xGChain),
      avg_diffopp = mean(diff_opponent),
    ) %>%
    arrange(desc(avg_avgpoints),
            sum_price)
)
```

Let's say the budget for the goalkeeper position is 10.

```{r def05}
(
  gk05 <- gk04 %>%
    arrange(desc(avg_avgpoints)) %>%
    filter(sum_price < 10 & sum_price > 8) %>%
    slice_head(n = 5) %>%
    left_join(gk03, by = "id_gk") %>%
    relocate(id_gk, player_team)
)
```

...it's going to be easier to plot the data.

```{r def_plot}
  (
    gk_plot01 <- gk05 %>%
      select(
        id_gk,
        sum_price,
        avg_avgpoints,
        avg_avgminutes,
        avg_avgict,
        avg_avginfluence,
        avg_avgXG,
        avg_avgxA,
        avg_avgxGChain,
        avg_diffopp
      ) %>%
      group_by(id_gk) %>%
      slice_head(n = 1) %>%
      ungroup() %>%
      pivot_longer(
        cols = 2:10,
        names_to = "indicator",
        values_to = "value"
      ) %>%
      ggplot(aes(
        x = reorder_within(id_gk, value, indicator),
        y = value,
        fill = factor(id_gk)
      )) +
      geom_col() +
      #geom_bar(stat = "identity", aes(fill = factor(id_gk))) +
      coord_flip() +
      scale_x_reordered() +
      #scale_fill_viridis(discrete = TRUE) +
      #scale_fill_manual(values = wes_palette("Darjeeling1")) +
      scale_fill_manual(values = c(
        "#ef476f", "#ffd166", "#06d6a0", "#118ab2", "#073b4c"
      )) +
      egg::theme_article() +
      facet_wrap(~ indicator, ncol = 3, scales = "free") +
      theme(legend.position = "none") +
      labs(x = "",
           y = "")
)
```

Which DEF team?

```{r def_choice}
gk_team <- "823"

(gk_choice <-
    gk05 %>% filter(id_gk == gk_team) %>% mutate("position" = "GK"))
```

Take a loot at the player's statistics individually.

```{r def_indv_stat}
(
  gk_plot02 <- gk_choice %>%
    select(
      player_team,
      now_cost,
      avg_points,
      avg_minutes,
      avg_ict,
      avg_influence,
      avg_xG,
      avg_xA,
      avg_xGChain,
      diff_opponent
    ) %>%
    pivot_longer(
      cols = 2:10,
      names_to = "indicator",
      values_to = "value"
    ) %>%
    ggplot(aes(
      x = reorder_within(player_team, value, indicator),
      y = value,
      fill = factor(player_team)
    )) +
    geom_col() +
    #geom_bar(stat = "identity", aes(fill = factor(id_gk))) +
    coord_flip() +
    scale_x_reordered() +
    #scale_fill_viridis(discrete = TRUE) +
    #scale_fill_manual(values = wes_palette("Darjeeling1")) +
    scale_fill_manual(values = c("#ef476f", "#ffd166")) +
    egg::theme_article() +
    facet_wrap( ~ indicator, ncol = 3, scales = "free") +
    theme(legend.position = "none") +
    labs(x = "",
         y = "")
)
```
